Lattice Boltzmann study of mass transfer for Bretherton/Taylor bubble train flow

### 2012-05-28 Text/introduction.tex
Final fixes.

Note that the parameters of the lattice
Boltzmann scheme are connected with  physical parameters only through  non-dimensional
numbers governing the physics of the problem. In our case, this number is the Peclet number, $Pe=\frac{\ububble L}{D}$.
Therefore, one can substitute any quantity, i.e.
$\ububble$ in the lattice Boltzmann units as soon as the
Peclet number is  matched in physical space and numerical simulations. This fact that $\ububble$
can be chosen arbitrarily is extremely useful in the context of numerical simulations, since it can
increase the time step, or decrease the computational demand by order of magnitude to obtain
meaningful results. This point will be used in simulations and covered later.

Next section will cover LBM benchmarks that resemble the mass transfer problem from a bubble, as 
mass transfer from the film  and circumferences.

\subsection{The radial case}
The case to be examined  is the mass transfer from the stair-case boundary forming a circle. It
can be
prescribed by the following system of equations:
\beq
\begin{aligned}
&\partial_t C(r,t)=\frac{1}{r}\partial_r r \partial_r C(r,t)\\
&C(a,t)=C_0,\,C(r,0)=C_{init}
\end{aligned}
\feq 
The solution is represented as \cite{chemical-correlations}:
\beq
\frac{C(r,t)-C_{0}}{C_{init}-C_{0}}=\sum_{n=1}^{\infty}{\frac{2}{\mu_n
J_1(\mu_n)}\exp\Bigl(-\mu_n^2 \frac{D t}{a^2} \Bigr)J_0\left(\mu_n \frac{r}{a}\right)},
\feq
where $\mu_n$ is the $n$-th zero root of the $0$th order Bessel polynomial $J_0(\mu_n)=0$. Some of
the corresponding roots are as follows: $\mu_1=2.4048$, $\mu_2=5.5201$, $\mu_3=8.6537$,
$\mu_4=11.7915$, $\mu_5=14.9309$.
By taking initial concentration to be $0$, one can obtain the following solution:
\beq
C(r,t)=C_0 \Biggl(1 - \sum_{n=1}^{\infty}{\frac{2}{\mu_n
J_1(\mu_n)}\exp\Bigl(-\mu_n^2 \frac{D t}{a^2} \Bigr)J_0(\mu_n \frac{r}{a})}\Biggr).
\feq

@@

\subsection{Parabolic velocity profile with the velocity zero gradient at the wall}
The problem is defined in terms of PDEs as follows: 
\beq
\begin{aligned}
&\frac{\partial C}{\partial x} U(y)=D\frac{\partial^2 C}{\partial y^2}\\
&C(0,y)=C_0,\, C(x,0)=\cstar,\, \frac{\partial C}{\partial y }(x,\delta)=0\\
&U(y)=U_0 \Bigl(\frac{y}{\delta}\Bigr)^2.
\end{aligned}
\feq
The procedure to solve this PDE is presented in Appendix \ref{appendix:zero:gradient}:
\beq
\label{analytical:solution:zero:gradient}
\begin{aligned}
&C(x,y)=\cstar+\sum_{m}{C_m
\sqrt{\frac{y}{\delta}}J_{\frac{1}{4}}\Bigl(\frac{m^2}{2}\frac{y^2}{\delta^2}\Bigr)\exp\Bigl(-\frac{
m^4 } { Pe }
\frac { x }{\delta}\Bigr)}\\
&C_m = (C_0-\cstar) \frac{\int_{\xi=0}^{1}{\xi^{5/2} J_{\frac{1}{4}}\Bigl(\frac{m^2
\xi^2}{2}\Bigr)\mathrm{d}\xi}}{\int_{\xi=0}^{1}{\xi^3 J_{\frac{1}{4}}^2\Bigl(\frac{m^2
\xi^2}{2}\Bigr)\mathrm{d}\xi}}
\end{aligned}
\feq
Fig. \ref{fig:parabolic:zero:gradient} shows the steady-state contours comparison between the
analytical solution and the simulation with the anti bounce-back boundary conditions used to impose constant
concentrations at the wall $\cstar=1$ and the inlet $C_0=0$. The simulation grid is
$80\times1600$. Parameter $\omegaminus=1.8$ implies the diffusion coefficient to be
$D=\frac{1}{3}\Bigl(\frac{1}{\omega}-\frac{1}{2}\Bigr)=0.0185$.  
\begin{figure}[htb!]
\includegraphics[width=\textwidth]{Figures/parabolic_profile_zero_gradient_comparison.eps}
\caption{Concentration contour levels comparison between the analytical solution, Eq.
\ref{analytical:solution:zero:gradient}, and simulation for the diffusion coefficient $D=0.0185$. Simulation was
performed on the grid size of $80\times 1600$. The constant concentration is imposed with the
anti bounce-back conditions, Eq. \ref{antibb}. Coordinate $x$ and $y$ are non-dimensionalized on
the height of the channel. $\delta=H/2$ is the half of the channel height.
\label{fig:parabolic:zero:gradient}}
\end{figure}

\subsection{Poiseulle velocity parabolic profile}
The problem we want to address can be formulated through the following PDE:
\beq
\begin{aligned}
&\frac{\partial C}{\partial x} U(y)=D\frac{\partial^2 C}{\partial y^2}\\
&C(0,y)=0,\, C(x,\pm \delta)=\cstar,\, \frac{\partial C}{\partial y }(x,0)=0\\
&U(y)=U_0 \Bigl(1-\bigl(\frac{y}{\delta}\bigr)^2\Bigr)
\end{aligned}
\feq
The procedure to solve this problem is presented in Appendix \ref{appendix:poiseuille} which yields
the final solution as:
\begin{equation}
C=\cstar-\cstar \sum_{m=0}{C_m e^{-m^4 \frac{x}{\delta}\frac{1}{Pe}} e^{-m^2
y^2/(2\delta^2)}{_1F_1}\Bigl(-\frac{m^2}{4}+\frac{1}{4},\frac{1}{2},m^2 \frac{y^2}{\delta^2}\Bigr)},
\end{equation}
where coefficients $C_m$ are taken from Eq. \ref{coeff:series:parabolic:profile}. The comparison
between contours of analytical and simulation results is presented in Fig.
\ref{fig:parabolic:comparison}. Parameters were taken  same as in the previous case: the
diffusion $D=0.0185$, the grid dimension is $80\times1600$. 
\begin{figure}[htb!]
\includegraphics[width=\textwidth]{Figures/parabolic_profile_comparison.eps}
\caption{Comparison between the analytical concentration contours and simulations with pressure
anti bounce-back conditions, Eq. \ref{antibb}. The parameters were
taken as $D=0.0185$ and grid $80\times1600$. \label{fig:parabolic:comparison}}
\end{figure}

@@

\section{Numerical approach}
\label{sec:numerics}
The multiphase code was utilized to obtain the flow patterns and bubble shapes for different capillary numbers
\cite{kuzmin-binary2d}. Five particular cases were chosen to be examined, their results are summarized
in Table \ref{table:capillary:cases}. 
\begin{table}[htb!]
\begin{tabularx}{\textwidth}{|X|X|X|X|X|X|X|X|X|}
\hline
$Ca$    &$Re$     &$\ububble$ &$\delta$&$\holdup$
&$\uliq$&$\ugas$&$\lbubble$&$\lslug$\\
\hline
$0.097$ &$1.656$  &$0.0055$ &$0.092$ &$0.30$ &$0.0046$&$0.0016$&$5.79$&$9.21$\\ 
$0.254$ &$4.318$  &$0.0143$ &$0.132$ &$0.28$ &$0.0108$&$0.0041$&$6.12$&$8.88$\\ 
$0.526$ &$8.938$  &$0.0297$ &$0.157$ &$0.27$ &$0.0209$&$0.0080$&$6.19$&$8.81$\\
$0.750$ &$12.744$ &$0.0424$ &$0.167$ &$0.25$ &$0.0293$&$0.0107$&$5.96$&$9.04$\\
$1.040$ &$17.665$ &$0.0588$ &$0.177$ &$0.22$ &$0.0397$&$0.0135$&$5.59$&$9.41$\\
\hline
\end{tabularx}
\caption{Sample results with the binary liquid lattice Boltzmann model \cite{kuzmin-binary2d}. The
following notations are used: the capillary number $Ca=\frac{\ububble L}{\rho \nu_{\mathrm{liq}}}$, $\delta$ is the
non-dimensional film thickness, $\uliq$ is the superficial liquid velocity, $\ugas$ is the
superficial gas velocity. The simulation sketch is presented in Fig.
\ref{fig:benchmark:hydro}. \label{table:capillary:cases}}
\end{table}
Note that velocities in Table \ref{table:capillary:cases} are small. It means that to
match large Peclet numbers, $Pe=\frac{\ububble L}{D}$, usually used in experiments, one needs to decrease the diffusion coefficient
$D=\frac{1}{3}\Bigl(\frac{1}{\omegaminus}-\frac{1}{2}\Bigr)$. Thus, parameter $\omegaminus\approx 0.5$. However, for such parameters
$\omegaminus$ the stability of the lattice Boltzmann method drastically decreases
\cite{kuzmin-d1q3}. On the other hand, one iteration in the lattice Boltzmann system corresponds
to the time step in the physical domain  $\Delta t=U_{\mathrm{bubble,LB}} \frac{\Delta
x}{U_{\mathrm{bubble,phys}}}$. One iteration time is proportional to the velocity $U_{\mathrm{LB}}$
and the typical number of simulation steps to obtain the steady-state mass transfer coefficient for
$Ca<0.2$ is of the order of a few millions. Therefore, it is desirable to increase  velocity  $U_{\mathrm{LB}}$
while keeping the Peclet number the same. If one increases velocity, then
$\omegaminus$ increases as well. Thus, the increase of the velocity impacts
positively on stability of the LBM as $\omegaminus$ is not close to its
stability limit $0.5$.

Given all the considerations above, mass transfer simulations are performed as:
\begin{description}
 \item[Flow field] Given the capillary number $Ca$, one needs to obtain hydrodynamic fields around
the bubble using the multiphase binary liquid lattice Boltzmann model according to work
\cite{kuzmin-binary2d}. Periodic boundary conditions are used. The grid used in this work is
$202\times 3000$ which corresponds to the fluid domain  of size $200\times3000$. The grid resolution is
taken to ensure grid
independency of results \cite{kuzmin-binary2d}. Note that we do not approximate bubble shapes by correlations, but we directly resolve them by using the multiphase solver. 
 \item[Bubble reference frame] Once the hydrodynamics is resolved, the mass transfer simulations
are conducted in the reference frame moving with the bubble. Then the bubble stands still and the
flow is  happening around the
bubble. We impose a steady concentration on the surface
of the bubble with the anti bounce-back condition, Eq. \ref{antibb}.
 \item[Velocity improvement] One can scale the velocity to
perform faster simulations. However, before doing it one needs to improve velocity. This issue arises because of the
multiphase model used in simulations. The binary liquid lattice Boltzmann model is the diffusive
interface model. Thus, the
transition between gas and liquid is the continuous function. We obtain the bubble shape according
to the phase indicator $\phi$ as used in \cite{kuzmin-binary2d}, say $\phi\leq0$. The velocity of the
bubble is calculated as the bubble tip velocity. Because of the square grid the shape of the bubble is determined within accuracy of one grid node. Thus,
there is the error in determination of the bubble velocity. Though these errors are small, however
there is a small non-zero velocity component pointing to the bubble in some places, see Fig. 8 in
\cite{kuzmin-binary2d} where some streamlines are penetrating a bubble surface a bit.
This small velocity is amplified upon the velocity scaling and is inconsistent with the
advection-diffusion equation leading to the instability after many iterations.

Thus, before performing the mass transfer simulations one additional single phase hydrodynamic
simulation is performed. An additional free-surface solver was developed in order to obtain velocity field
consistency with the advection-diffusion equation. We take results from the
multiphase simulations, extract a
bubble shape using the phase indicator $\phi\leq0$, approximate the bubble shape by the stair-case
line. Then, the obtained bubble velocity is imposed on walls. This corresponds to conducting
simulation in the reference frame moving with the bubble. In addition, the free-slip boundary
condition on the bubble surface is imposed. Appendix \ref{appendix-free-surface} covers the simple
free-slip boundary conditions implementation drastically improving velocity patterns. The system is
iterated until a
steady state is reached. Note, that these type of simulations are much faster than original
multiphase simulations. As the output all the non-zero velocity components perpendicular to the
bubble surface are completely eliminated. We compared original multiphase simulations with
one-component free-slip simulations. All quantities as superficial slug and liquid velocities are
within $3\%$ for the capillary number in the range $0.05\leq Ca \leq 1.0$. This simulation helps to
make mass transfer simulations stable and faster.%, since the velocity scaling can be done easier and
%streamlines are non-penetrating through the interface. 
One
can see in Fig. \ref{fig:free:surface} two streamlines profiles for $Ca=0.097$  and $Ca=1.040$.
\begin{figure}[htb!]
\includegraphics[angle=90,width=\textwidth]{Figures/performed_ca0097_out.eps}\\
\includegraphics[angle=90,width=\textwidth]{Figures/performed_ca1040_out.eps}
\caption{The streamlines patterns produced by the free-surface solver with simplistic approximation
of the free-slip bubble surface, see Appendix \ref{appendix-free-surface}. Two completely different
velocity patterns are covered, $Ca=0.097$ (top) and $Ca=1.040$ (bottom).
\label{fig:streamlines:tweaked:velocity}}
\end{figure}

@@

\section{Results}
This section covers simulation results. We first examine the possibility to increase the fluid
velocity while keeping the Peclet number the same. After the periodic boundary
conditions for $5$ capillary number cases will be presented. Finally, we will examine
many
cells simulations for two representative cases velocity patterns, as $Ca=0.0907$ and $Ca=1.04$
(see Fig. \ref{fig:streamlines:tweaked:velocity}).

\subsection{Keeping the Peclet number}

@@

This section addresses the simulations performance. The ultimate goal is to increase
velocity magnitude significantly while keeping dimensionless parameters the same to

@@

simulations. This is especially important to be able to simulate a few unit cells simulations. For example, ten unit cells simulations require a grid of $30000\times202$ nodes.

@@

As it was discussed above, only the Peclet number governs 
the advection-diffusion equation. In the lattice Boltzmann system it is represented as:

@@

As far as we want to increase velocity, one needs to increase the diffusion coefficient as well. The runs were made with

@@

its corresponding capillary numbers are presented in Table \ref{table:scaling:peclet}. Periodic

@@

\caption{This table presents the achievable stable velocity $\ububble$ when one scales velocity. One
iteration in the lattice
Boltzmann system corresponds to time difference as
$\Delta t=\frac{U_{\mathrm{bubble,LB}}}{U_{\mathrm{bubble,phys}}} \Delta x$. Therefore the speed up
of the simulation is proportional to the  $U_{\mathrm{bubble,LB}}$. Note that time
iterations
indicated in the table correspond to the one time moment in the physical space. One can see that
scaling produces adequate results.  The achievable stable velocity is around $U_{bubble}=0.1$. Average concentrations are in the reasonable agreement. The contour

@@

The corresponding to Table \ref{table:scaling:peclet} concentration contour profiles for different

@@

acceptable agreement between all of them. Note, that the speedup can be upto $10$ to $40$ times
faster.

@@

\subsection{Average concentration}
\label{main:results:periodic}
As indicated in Section \ref{section:cases} one can calculate the average domain concentration over
the time. Eq. \ref{theor:one:concentration:time} shows how 
the volumetric mass transfer coefficient depends on
time, i.e. streamwise coordinate:
\beqal
&\vol t \frac{\ububble}{\ugas+\uliq}=\ln\frac{\cstar}{\cstar-\langle C(t) \rangle}\\
&\vol \frac{\lunit}{\ugas+\uliq}=\frac{\lunit}{\ububble t} \ln \frac{C^*}{C^*-\langle C(t)
\rangle},\\
\feqal
where $\langle C(t) \rangle$ is the average concentration in the liquid domain.
Simulations for coefficient $\vol \frac{\ububble}{\ugas+\uliq}$ are shown in Fig.
\ref{fig:aver:conc:different:capillaries:time} for different Peclet numbers and velocity scalings
indicated in Table \ref{table:scaling:peclet}. When the average concentration goes
close to $\cstar=1$ then Eq. \ref{theor:one:concentration:time} gives inadequate results due to
 accuracy of the logarithmic function. Fig. \ref{fig:aver:conc:different:capillaries:time} presents
the volumetric mass transfer coefficient against number of iterations. However, due to scaling
each simulation has a different physical time step. We scaled curves against number of cell
units which tracer will pass with the given bubble velocity, i.e. 
$N_{\text{cell\,units}}=\frac{\text{scale}\cdot \ububble\cdot N_{\mathrm{iter}}}{\lunit}$.  Fig.
\ref{fig:aver:conc:different:capillaries} shows the volumetric mass transfer
dependency against the distance in unit cells length. One can see in Table
\ref{table:steady:state:average} that for different Peclet numbers different time (number of unit
cells) is required to achieve the steady state. For example, for
larger Peclet number less number of iterations are required to achieve the steady state condition.

>>>>

Note that the parameters of the lattice
Boltzmann scheme are connected with  physical parameters only through  non-dimensional
numbers governing the physics of the problem. In our case, this number is the Peclet number, $Pe=\frac{\ububble L}{D}$.
Therefore, one can substitute any quantity, i.e.
$\ububble$ in the lattice Boltzmann units as soon as the
Peclet number is  matched in physical space and numerical simulations. The fact that $\ububble$
can be chosen arbitrarily is extremely useful in the context of numerical simulations, since one can
increase the time step and decrease the computational demand by an order of magnitude while still
obtaining meaningful results. This point will be used in simulations and covered later.

% TODO: I'm not sure I understand this sentence.  Does it say that we use the cap + film simplification
% for the simulations, or that the simulation results are compared to results obtained via this simplification?
Next section will cover LBM benchmarks that resemble the mass transfer problem from a bubble, as 
mass transfer from the film  and circumferences.

\subsection{The radial case}
The case to be examined  is the mass transfer from a stair-case boundary forming a circle. It
can be described by the following system of equations:
\beq
\begin{aligned}
&\partial_t C(r,t)=\frac{1}{r}\partial_r r \partial_r C(r,t)\\
&C(a,t)=C_0,\,C(r,0)=C_{init}
\end{aligned}
\feq 
The solution is represented as \cite{chemical-correlations}:
\beq
\frac{C(r,t)-C_{0}}{C_{init}-C_{0}}=\sum_{n=1}^{\infty}{\frac{2}{\mu_n
J_1(\mu_n)}\exp\Bigl(-\mu_n^2 \frac{D t}{a^2} \Bigr)J_0\left(\mu_n \frac{r}{a}\right)},
\feq
where $\mu_n$ is the $n$-th zero root of the $0$th order Bessel polynomial $J_0(\mu_n)=0$. Some of
the corresponding roots are as follows: $\mu_1=2.4048$, $\mu_2=5.5201$, $\mu_3=8.6537$,
$\mu_4=11.7915$, $\mu_5=14.9309$.
By taking initial concentration to be $0$, one can obtain the following solution:
\beq
C(r,t)=C_0 \Biggl(1 - \sum_{n=1}^{\infty}{\frac{2}{\mu_n
J_1(\mu_n)}\exp\Bigl(-\mu_n^2 \frac{D t}{a^2} \Bigr)J_0(\mu_n \frac{r}{a})}\Biggr).
\feq

@@

\subsection{Parabolic velocity profile with the velocity zero gradient at the wall}
The problem is defined in terms of PDEs as follows: 
\beq
\begin{aligned}
&\frac{\partial C}{\partial x} U(y)=D\frac{\partial^2 C}{\partial y^2}\\
&C(0,y)=C_0,\, C(x,0)=\cstar,\, \frac{\partial C}{\partial y }(x,\delta)=0\\
&U(y)=U_0 \Bigl(\frac{y}{\delta}\Bigr)^2.
\end{aligned}
\feq
The procedure to solve this PDE is presented in Appendix \ref{appendix:zero:gradient}:
\beq
\label{analytical:solution:zero:gradient}
\begin{aligned}
&C(x,y)=\cstar+\sum_{m}{C_m
\sqrt{\frac{y}{\delta}}J_{\frac{1}{4}}\Bigl(\frac{m^2}{2}\frac{y^2}{\delta^2}\Bigr)\exp\Bigl(-\frac{
m^4 } { Pe }
\frac { x }{\delta}\Bigr)}\\
&C_m = (C_0-\cstar) \frac{\int_{\xi=0}^{1}{\xi^{5/2} J_{\frac{1}{4}}\Bigl(\frac{m^2
\xi^2}{2}\Bigr)\mathrm{d}\xi}}{\int_{\xi=0}^{1}{\xi^3 J_{\frac{1}{4}}^2\Bigl(\frac{m^2
\xi^2}{2}\Bigr)\mathrm{d}\xi}}
\end{aligned}
\feq
Fig. \ref{fig:parabolic:zero:gradient} shows a comparison of the steady-state contours between the
analytical solution and the simulation with anti bounce-back boundary conditions used to impose constant
concentrations at the wall $\cstar=1$ and the inlet $C_0=0$. The simulation grid is
$80\times1600$. Parameter $\omegaminus=1.8$ implies the diffusion coefficient to be
$D=\frac{1}{3}\Bigl(\frac{1}{\omega}-\frac{1}{2}\Bigr)=0.0185$.  
\begin{figure}[htb!]
\includegraphics[width=\textwidth]{Figures/parabolic_profile_zero_gradient_comparison.eps}
\caption{Comparison of the concentration contour levels between the analytical solution, Eq.
\ref{analytical:solution:zero:gradient}, and simulation for the diffusion coefficient $D=0.0185$. The simulation was
performed on the grid size of $80\times 1600$. The constant concentration is imposed with the
anti bounce-back conditions, Eq. \ref{antibb}. Coordinate $x$ and $y$ are non-dimensionalized on
the height of the channel. $\delta=H/2$ is half of the channel height.
\label{fig:parabolic:zero:gradient}}
\end{figure}

\subsection{Poiseulle velocity parabolic profile}
The problem we want to address can be formulated through the following PDE:
\beq
\begin{aligned}
&\frac{\partial C}{\partial x} U(y)=D\frac{\partial^2 C}{\partial y^2}\\
&C(0,y)=0,\, C(x,\pm \delta)=\cstar,\, \frac{\partial C}{\partial y }(x,0)=0\\
&U(y)=U_0 \Bigl(1-\bigl(\frac{y}{\delta}\bigr)^2\Bigr)
\end{aligned}
\feq
The procedure to solve this problem is presented in Appendix \ref{appendix:poiseuille} which yields
the final solution as:
\begin{equation}
C=\cstar-\cstar \sum_{m=0}{C_m e^{-m^4 \frac{x}{\delta}\frac{1}{Pe}} e^{-m^2
y^2/(2\delta^2)}{_1F_1}\Bigl(-\frac{m^2}{4}+\frac{1}{4},\frac{1}{2},m^2 \frac{y^2}{\delta^2}\Bigr)},
\end{equation}
where coefficients $C_m$ are taken from Eq. \ref{coeff:series:parabolic:profile}. The comparison
between contours of analytical and simulation results is presented in Fig.
\ref{fig:parabolic:comparison}. Parameters were taken  same as in the previous case: the
diffusion coefficient is $D=0.0185$, the grid dimension is $80\times1600$. 
\begin{figure}[htb!]
\includegraphics[width=\textwidth]{Figures/parabolic_profile_comparison.eps}
\caption{Comparison between the analytical concentration contours and simulations with pressure
anti bounce-back conditions, Eq. \ref{antibb}. The simulation was done for $D=0.0185$ with a
$80\times1600$ grid. \label{fig:parabolic:comparison}}
\end{figure}

@@

\section{Numerical approach}
\label{sec:numerics}
The multiphase code was utilized to obtain the flow patterns and bubble shapes for different capillary numbers
\cite{kuzmin-binary2d}. Five particular cases were chosen to be examined, their results are summarized
in Table \ref{table:capillary:cases}. 
\begin{table}[htb!]
\begin{tabularx}{\textwidth}{|X|X|X|X|X|X|X|X|X|}
\hline
$Ca$    &$Re$     &$\ububble$ &$\delta$&$\holdup$
&$\uliq$&$\ugas$&$\lbubble$&$\lslug$\\
\hline
$0.097$ &$1.656$  &$0.0055$ &$0.092$ &$0.30$ &$0.0046$&$0.0016$&$5.79$&$9.21$\\ 
$0.254$ &$4.318$  &$0.0143$ &$0.132$ &$0.28$ &$0.0108$&$0.0041$&$6.12$&$8.88$\\ 
$0.526$ &$8.938$  &$0.0297$ &$0.157$ &$0.27$ &$0.0209$&$0.0080$&$6.19$&$8.81$\\
$0.750$ &$12.744$ &$0.0424$ &$0.167$ &$0.25$ &$0.0293$&$0.0107$&$5.96$&$9.04$\\
$1.040$ &$17.665$ &$0.0588$ &$0.177$ &$0.22$ &$0.0397$&$0.0135$&$5.59$&$9.41$\\
\hline
\end{tabularx}
\caption{Sample results with the binary liquid lattice Boltzmann model \cite{kuzmin-binary2d}. The
following notations are used: the capillary number $Ca=\frac{\ububble L}{\rho \nu_{\mathrm{liq}}}$, $\delta$ is the
non-dimensional film thickness, $\uliq$ is the superficial liquid velocity, $\ugas$ is the
superficial gas velocity. The simulation sketch is presented in Fig.
\ref{fig:benchmark:hydro}. \label{table:capillary:cases}}
\end{table}
Note that velocities in Table \ref{table:capillary:cases} are small. It means that to
match large Peclet numbers, $Pe=\frac{\ububble L}{D}$, usually used in experiments, one needs to decrease the diffusion coefficient
$D=\frac{1}{3}\Bigl(\frac{1}{\omegaminus}-\frac{1}{2}\Bigr)$. Thus, parameter $\omegaminus\approx 0.5$. However, for such parameters
$\omegaminus$ the stability of the lattice Boltzmann method drastically decreases
\cite{kuzmin-d1q3}. On the other hand, one iteration in the lattice Boltzmann system corresponds
to the physical time step  $\Delta t=U_{\mathrm{bubble,LB}} \frac{\Delta
x}{U_{\mathrm{bubble,phys}}}$. The iteration time is proportional to the velocity $U_{\mathrm{LB}}$
and the typical number of simulation steps to obtain the steady-state mass transfer coefficient for
$Ca<0.2$ is of the order of a few millions. Therefore, it is desirable to increase  velocity  $U_{\mathrm{LB}}$
while keeping the Peclet number the same. If one increases the velocity, then
$\omegaminus$ increases as well, moving away from the stability limit $0.5$ and therefore
stabilizing the simulation.

Given all the considerations above, mass transfer simulations are performed as follows:
\begin{description}
 \item[Flow field] Given the capillary number $Ca$, one needs to obtain hydrodynamic fields around
the bubble using the multiphase binary liquid lattice Boltzmann model according to work
\cite{kuzmin-binary2d}. Periodic boundary conditions are used. The grid used in this work is
$202\times 3000$ which corresponds to the fluid domain  of size $200\times3000$. The grid resolution is
taken to ensure grid
independency of the results \cite{kuzmin-binary2d}. Note that we do not approximate bubble shapes by correlations, but we directly resolve them by using the multiphase solver. 
 \item[Bubble reference frame] Once the hydrodynamics is resolved, the mass transfer simulations
are conducted in the reference frame moving with the bubble, where the bubble stands still and the
flow is happening around the
bubble. We impose a steady concentration on the surface
of the bubble with the anti bounce-back condition, Eq. \ref{antibb}.
 \item[Velocity improvement] One can scale the velocity to
perform faster simulations. However, before doing it one needs to improve the velocity field.
This issue arises because of the
multiphase model used in simulations. The binary liquid lattice Boltzmann model is a diffusive
interface model where no clear boundary between gas and liquid exists.
% TODO: Is this only a problem of how we determine the bubble shape or also a problem of
% spurious velocities within the model itself?
We obtain the bubble shape by imposing a condition on the order parameter field $\phi$ as in \cite{kuzmin-binary2d}, say $\phi\leq0$. The velocity of the
bubble is calculated as the bubble tip velocity. Because of the square grid the shape of the bubble is determined within accuracy of one grid node. Thus,
there is the error in determination of the bubble velocity. Though these errors are small,
there is still a small non-zero velocity component pointing to the bubble in some places, see Fig. 8 in
\cite{kuzmin-binary2d} where some streamlines are penetrating the bubble surface a bit.
This small velocity is amplified upon the velocity scaling and is inconsistent with the
advection-diffusion equation leading to the instability after many iterations.

Thus, before performing the mass transfer simulations one additional single phase hydrodynamic
simulation is performed. An additional free-surface solver was developed in order to obtain velocity field
consistent with the advection-diffusion equation. We take results from the
multiphase simulations, extract a
bubble shape using the phase indicator $\phi\leq0$, and approximate the bubble shape by the stair-case
line. The obtained bubble velocity is then imposed on the walls. This corresponds to conducting the
simulation in the reference frame moving with the bubble. In addition, the free-slip boundary
condition on the bubble surface is imposed. Appendix \ref{appendix-free-surface} covers the simple
free-slip boundary condition implementation drastically improving velocity patterns. The system is
iterated until a
steady state is reached. Note, that these type of simulations are much faster than original
multiphase simulations. As the output all the non-zero velocity components perpendicular to the
bubble surface are completely eliminated. We compared original multiphase simulations with
one-component free-slip simulations. All quantities such as superficial slug and liquid velocities are
within $3\%$ for the capillary number in the range $0.05\leq Ca \leq 1.0$. 
One can see in Fig. \ref{fig:free:surface} two streamline profiles for $Ca=0.097$  and $Ca=1.040$.
\begin{figure}[htb!]
\includegraphics[angle=90,width=\textwidth]{Figures/performed_ca0097_out.eps}\\
\includegraphics[angle=90,width=\textwidth]{Figures/performed_ca1040_out.eps}
\caption{The streamlines patterns produced by the free-surface solver with simplistic approximation
of the free-slip bubble surface, see Appendix \ref{appendix-free-surface}. Two completely different
velocity patterns are covered, $Ca=0.097$ (top) and $Ca=1.040$ (bottom).
\label{fig:streamlines:tweaked:velocity}}
\end{figure}

@@

\section{Results}
This section covers simulation results. We first examine the possibility to increase the fluid
velocity while keeping the Peclet number the same. After the periodic boundary
conditions for $5$ capillary number cases will be presented. Finally, we will examine
many
cell simulations for two representative cases of velocity patterns, $Ca=0.0907$ and $Ca=1.04$
(see Fig. \ref{fig:streamlines:tweaked:velocity}).

\subsection{Keeping the Peclet number constant}

@@

This section addresses the process of significantly increasing the
velocity magnitude while keeping dimensionless parameters the same to

@@

simulations. This is especially important to be able to simulate a few unit cells in a reasonable time.
For example, ten unit cell simulations require a grid of $30000\times202$ nodes.

@@

Since the Peclet number is the only dimensionless quantity governing
the advection-diffusion equation:

@@

one needs to increase the diffusion coefficient when the velocity is increased. Simulation runs were made with

@@

their corresponding capillary numbers are presented in Table \ref{table:scaling:peclet}. Periodic

@@

\caption{This table presents the achievable stable velocity $\ububble$ when one scales velocity. Since
the physical time step represented by a single iteration of the simulation is directly proportional
$U_{\mathrm{bubble,LB}}$, scaling the velocity directly translates to an effective speed-up of the simulation.
Note that time iterations
indicated in the table correspond to physical time. One can see that
scaling produces adequate results.  The achievable stable velocity is around $U_{bubble}=0.1$. Average concentrations are in reasonable agreement. The contour

@@

The concentration contour profiles corresponding to Table \ref{table:scaling:peclet} for different

@@

acceptable agreement between all of them. Note, that the speedup can be up to $10$ to $40$ times
higher.

@@

\subsection{Average concentration}
\label{main:results:periodic}
As indicated in Section \ref{section:cases} one can calculate the average domain concentration over
the time. Eq. \ref{theor:one:concentration:time} shows how 
the volumetric mass transfer coefficient depends on
time, i.e. streamwise coordinate:
\beqal
&\vol t \frac{\ububble}{\ugas+\uliq}=\ln\frac{\cstar}{\cstar-\langle C(t) \rangle}\\
&\vol \frac{\lunit}{\ugas+\uliq}=\frac{\lunit}{\ububble t} \ln \frac{C^*}{C^*-\langle C(t)
\rangle},\\
\feqal
where $\langle C(t) \rangle$ is the average concentration in the liquid domain.
Simulations for coefficient $\vol \frac{\ububble}{\ugas+\uliq}$ are shown in Fig.
\ref{fig:aver:conc:different:capillaries:time} for different Peclet numbers and velocity scalings
indicated in Table \ref{table:scaling:peclet}. When the average concentration goes
close to $\cstar=1$ then Eq. \ref{theor:one:concentration:time} gives inadequate results due to
 accuracy of the logarithmic function. Fig. \ref{fig:aver:conc:different:capillaries:time} presents
the volumetric mass transfer coefficient against number of iterations. However, due to scaling
each simulation has a different physical time step. We scaled curves against number of cell
units which tracer will pass with the given bubble velocity, i.e. 
$N_{\text{cell\,units}}=\frac{\text{scale}\cdot \ububble\cdot N_{\mathrm{iter}}}{\lunit}$.  Fig.
\ref{fig:aver:conc:different:capillaries} shows the volumetric mass transfer
dependency against the distance in unit cell length. One can see in Table
\ref{table:steady:state:average} that for different Peclet numbers different time (number of unit
cells) is required to achieve the steady state. For example, for
larger Peclet numbers fewer iterations are required to achieve the steady state condition.
